---
title: "Assignment 9 - K-means"
author: "Md Ariful Islam Sohag"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Goal of this assignment  
The goal of this assignment is for you to **explore the k-means algorithm** learned in class.  

In specific, you will be asked to:  
  - Use a different number of k than what we did in class,  
  - Validate the results of the new model, and  
  - Compare your results with those found in class  

# Instructions  
You will first need to watch (and code along with) the following k-means video:

  - k-means lab: https://youtu.be/GFgMp5tYiMU?si=CI0E-2r-cYZLGVL1 (start from beginning of video, watch till the 01:10)  
  - The partial script for the video above is in our GitHub repository, 04-classcode, **03-25_multivar_kmeans_partial.qmd**. Move the script mentioned above into the `code` subfolder of your `10_multivar` project
  
**AFTER watching the video and developing the code in class**, then start working on this HW assignment (instructions below):

  - Move this script into the `code` subfolder of your `10_multivar` project  
  - On this script, fill your name under the `title` section of the YAML  
  - Go back to the class k-means script, choose a different number of k than what we did in class, and run it.    

# Questions  
## 1. What number of k did you select? Why?  

# Setup  
```{r}
#| message: false
#| warning: false

#install.packages("ggcorrplot")
#install.packages("factoextra")
#install.packages("ggpmisc")

library(tidyverse)
library(ggcorrplot)
library(broom)
library(car)
library(factoextra)
library(ggpmisc)
```


```{r}
weather <- read_csv("../../09_opendata/data/weather_monthsum.csv")

weather
```

k-means does not normalize our data for us like PCA did, so we will need to do that before running the model. Therefore, the in the following chunk, I normalize our data.

```{r}
# normalizing the data
weather_norm <- weather_n %>%
  mutate(across(everything(), ~scale(.x) ))


weather_norm

summary(weather_norm)
```

Since the choice of k can be subjective, we will need to find an objective way to select the value of k that most properly represents our dataset. 

```{r choosing k - total error}
# Total error x k
fviz_nbclust(weather_norm,
             method = "wss",
             k.max = 10,
             FUNcluster = kmeans
             )
```
The above plot is known as total within sum square "wss" or elbow method. The principle of this method is that we will chose a number of k based on when we see an elbow. Based on our data set, the elbow will be within 3 to 5. 

```{r choosing k - silhouette}
# Silhouette width
fviz_nbclust(weather_norm, 
             method = "s",
             k.max = 10,
             FUNcluster = kmeans) 

```

Another method to chose the number of K is the silhouette width. According to the silhouette width, our data have the highest average silhouette width when cluster number is 4. Therefore, based on this method the number of cluster should be 4.

However, since I have to select a different number for the assignment instead of 4 I am selecting 3.

The reason behind to chose k=3 is that based on the elbow method and Silhouette width method, our data should have 3 to 5 cluster for best fit. Therforte, after k=4, the best choice for the number of cluster is k = 3.

```{r mod_km3 }
mod_km3 <- kmeans(weather_norm,
                  centers = 3,
                  nstart = 10
                  )

mod_km3
```


## 2. How many observations are there in each cluster?  

```{r}
weather %>%
  mutate(cluster = mod_km3$cluster) %>%
  group_by(cluster) %>%
  tally()
```

Therefore, cluster 1 have 108 observations, cluster 2 have 253 observations, and cluster 3 have 337 observations.

## 3. Using the `fviz_cluster()` function, import here the original plot with k=4 did in class, and the new plot with the number of k you selected for this exercise. How do they visually compare? Which one seems to be a better choice, and why?  

Before comparing the plots for k =4 and k = 3, I must import the km4 object in the following chunk
```{r mod_km4 }
mod_km4 <- kmeans(weather_norm,
                  centers = 4,
                  nstart = 10
                  )

mod_km4
```




```{r kmeans PCA plot for km3}
fviz_cluster(mod_km3,
             data = weather_norm
             )
```
when we have 3 cluster for our data, cluster 2 is skewed to the left and have a odd shape.

```{r kmeans PCA plot for km4}
fviz_cluster(mod_km4,
             data = weather_norm
             )
```

When we have 4 clusters for our data, cluster 1 have small number of observations but visually looking better than the plot when we have 3 clusters.

According to the elbow and Silhouette width method, our data should have 4 clusters. Visual plots also reflect that our data should have 4 clisters for the best explanation of our data.

# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-09-kmeans-LASTNAME.html`.    

For ex., mine would be `Assignment-09-kmeans-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assignment #9 - K-means** by April 3rd 11:59 pm.  
  




